\documentclass[../AlgebraQualSolutions.tex]{subfiles}

\begin{document}
\section{Unfinished}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	\subsection{Fall 2012}

	\begin{prob}{F12}{F12.G2}
	Define the center of a group.
	\begin{enumerate}[(a)]
	\item Prove that if the order of $G$ is $p^k$ for some prime $p$ then $G$ has nontrivial center.
	\item Suppose that $p$ and $q$ are distinct primes. Prove that a non-abelian group of order $pq$ has trivial center.
	\end{enumerate}
	\end{prob}
	
	The center of a group $G$ is the subgroup
		\[Z(G) = \{g \in G: gx = xg \textrm{ for any $x \in G$} \} \]
	
	\begin{proof}
	Assume that $|G| = p^k$ for some prime $p$. Suppose that $Z(G)$ is trivial and consists only of the identity element $e \in G$. If $k = 1$, the order of $G$ is prime which implies that $G$ is cyclic and therefore abelian. Therefore, assume that $k > 1$.
	\end{proof}
	
	\begin{prob}{F12}{F12.RF1}
	\begin{enumerate}[(a)]
	\item Let $\nu$ be a primitive 10th root of unity. Compute the Galois group of $\Q(\nu)$.
	\item let $K_t = \Q(\omega)$ where $\omega$ is a primitive $t$th root of unity. Show that $\sqrt[3]{2}$ is not in any $K_t$.
	\end{enumerate}
	\end{prob}
	
	\begin{prob}{F12}{F12.RF2}
	Suppose that $R,S,A$ are all commutative rings with unity.
	\begin{enumerate}[(a)]
	\item Suppose that $R$ is a PID, $S$ is an integral domain, and $\varphi:R \to S$ is a surjection. Prove that either $\varphi$ is an isomorphism or $S$ is a field.
	\item Prove that $A[x]$ is a PID if and only if $A$ is a field.
	\end{enumerate}
	\end{prob}
	
	\begin{prob}{F12}{F12.RF3}
	\begin{enumerate}[(a)]
	\item Prove that $\Z/m\Z$ has no non-zero nilpotent elements if and only if $m$ has no multiple prime factor.
	\item Prove that every element of $\Z/m\Z$ is either nilpotent or a unit whenever $m$ is a prime power.
	\item Prove that if $r$ is a nilpotent element of a ring with unity then $1-r$ is a unit.
	\end{enumerate}
	\end{prob}

	\begin{proof}
		Assume that $r \in R$ is a nilpotent element of a ring with unity. Then there exists $n \in \N$ such that $r^n = 0$. Then,
			\[1 = 1 - r^n = (1 + r + r^2 + \cdots + r^{n-1})(1-r)\]
		meaning that $\sum_{i=0}^{n-1}r^i$ is the multiplicative inverse of $(1-r)$. That is, $1-r$ is a unit.
	\end{proof}
	
	\begin{prob}{F12}{F12.LA1}
	Let $V$ be the vector space of real $n \times n$ matrices. Show that 	
		\[\langle A, B \rangle = n \tr(AB) - \tr(A)\tr(B) \]
	defines a symmetric bilinear form on $V$.
	\begin{enumerate}[(a)]
	\item Prove that $\langle,\rangle$ is singular.
	\item Prove that the restriction of $\langle,\rangle$ to the subspace $W$ of symmetric matrices with 0 trace is positive definite.
	\end{enumerate}
	\end{prob}

	\begin{prob}{F12}{F12.LA3}
	Let $M$ be a $5 \times 5$ matrix with rational entries whose characteristic polynomial is of the form
		\[(x^2+1)(x^3+x+q).\]
	For which, if any, $q \in \Q$ is it possible that there does not exist an invertible matrix $A$ with rational entries such that $A\inv M A$ has block diagonal form with a $2 \times 2$ block and a $3 \times 3$ block?
	\end{prob}
	
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\subsection{Spring 2012}


	\begin{prob}{S12}{S12.G1}
		Prove that there are at least two non-isomorphic non-abelian groups of order 24.
	\end{prob}

	\begin{prob}{S12}{S12.G2}
		\begin{enumerate}[(a)]
			\item State (any version of) the Fundamental Theorem of finite abelian groups.
			\item Classify all abelian groups of order 144.
			\item Explain which group in part (b) is isomorphic to the group $\Z_4 \times \Z_{36}$.
		\end{enumerate}
	\end{prob}

	\begin{prob}{S12}{S12.G3}
		\begin{enumerate}[(a)]
			\item State the Sylow Theorems.
			\item Prove that every group of order 126 has a normal subgroup of order 7.
			\item Prove that any group of order 1000 is not simple.
		\end{enumerate}
	\end{prob}

	\begin{prob}{S12}{S12.RF1}
		Prove or disprove:

		\begin{enumerate}[(a)]
			\item If $R$ is an integral domain, then $R[x]$ is an integral domain.
			\item If $R$ is a principal ideal domain, then $R[x]$ is a principal ideal domain.
		\end{enumerate}
	\end{prob}

	\begin{prob}{S12}{S12.RF2}
		Let $F$ be a finite field.

		\begin{enumerate}[(a)]
			\item Show that the multiplicative group $F^*$ is cyclic.
			\item Suppose $|F| = 125 = 5^3$ and $\cyc{\alpha} = F^*$. What is $\alpha^{62}$?
			\item Is there a $\beta \neq \alpha$ in $F$ such that $\cyc{\beta} = F^*$?
		\end{enumerate}
	\end{prob}

	\begin{prob}{S12}{S12.RF3}
		Let $\alpha = \sqrt{-1+\sqrt{2}}$.
		\begin{enumerate}[(a)]
			\item Prove that $\alpha$ is the root of a monic polynomial in $\Q[x]$.
			\item Let $K$ be the smallest Galois extension of $\Q$ that contains $\alpha$. Find the degree $[K: \Q]$.
		\end{enumerate}
	\end{prob}

	\begin{prob}{S12}{S12.LA1}

		Consider a real vector space $V = \R^n$ with the Euclidean inner product and let $U$ be a subspace of $V$.

		\begin{enumerate}[(a)]
			\item Prove that $U$ has an orthonormal basis. Note that this is the real version of the Gram-Schmidt Theorem.
			\item Find an orthonormal basis for the span of $(1,2,0)$ and $(1,1,3)$ in $\R^3$.
		\end{enumerate}
	\end{prob}

	\begin{prob}{S12}{S12.LA2}
		Let $V$ be a finite dimensional vector space. A linear transformation $T: V \to V$ is a projection when $T = T^2$. Prove that there exists a basis for $V$ such that the matrix for $T$ with respect to this basis is a diagonal matrix with diagonal entries all zeros or ones.
	\end{prob}

	\begin{prob}{S12}{S12.LA3}
		Let $V$ be a finite dimensional vector space over $\C$.

		\begin{enumerate}[(a)]
			\item Define the characteristic polynomial of a linear transformation of $V$ and the minimal polynomial of a linear transformation of $V$.
			\item Give an example of two lienar transformations $S,T: V \to V$ such that $S$ and $T$ have the same characteristic polynomial, but are not similar.
			\item Give an example of two linear transformations $S,T: V \to V$ such that $S$ and $T$ have the same minimal polynomial, but are not similar.
		\end{enumerate}
	\end{prob}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\subsection{Fall 2013}

	\begin{prob}{F13}{F13.G1}
		A subgroup $H$ of a group $G$ is \emph{characteristic} if $\alpha(H) = H$ for any automorphism $\alpha$ of $H$.
		
		\begin{enumerate}[(a)]
			\item Prove that if $H$ is characteristic in $K$ and $K$ is characteristic in $G$, then $H$ is characteristic in $G$.
			\item Suppose now that $H$, $K$, and $G$ are groups with $H$ a normal subgroup of $K$ and $K$ a normal subgroup of $G$. Does this imply that $H$ is a normal subgroup of $G$?
		\end{enumerate}
	\end{prob}

	\begin{prob}{F13}{F13.G2}
		Prove that every finite group of order greater than two has a non-trivial automorphism.
	\end{prob}

	\begin{prob}{F13}{F13.G3}

		\begin{enumerate}[(a)]
			\item Let $G$ be a finite group and $H$ a subgroup of $G$. Prove that the order of $H$ divides the order of $G$.
			\item  Let $\Q$ and $\Z$ denote the additive groups of the rationals and integers, respectively. Prove that $\Q/\Z$ has no proper subgroups of finite index.
		\end{enumerate}
	\end{prob}

	\begin{prob}{F13}{F13.RF1}
		Let $V = \R^2$, regarded as a two-dimensional subspace over $\R$. Let $L(V)$ be the ring of all linear transformations from $V$ to $V$. let $T \in L(V)$ be given by $T(x,y) = (y,-x)$ and define 
			\[A = \{S \in L(V): ST = TS\}. \]
		\begin{enumerate}[(a)]
			\item Prove that $A$ is a subring of $L(V)$.
			\item To what well-known ring is $A$ isomorphic?
		\end{enumerate}
	\end{prob}

	\begin{prob}{F13}{F13.RF2}

		\begin{enumerate}[(a)]
			\item Let $N$ be a non-negative integer and $\alpha \in \C$ a primitive $N$th root of unity. FOr which $N$ is it true that $\Q(\alpha) = (\alpha + \alpha\inv)$?
			\item Let $K$ be a field and $\beta$ an element of the algebraic closure of $K$. If $[K(3):K]$ is odd, prove that $K(\beta) = K(\beta + \beta\inv)$.
		\end{enumerate}
	\end{prob}

	\begin{prob}{F13}{F13.RF3}
		Let $F$ be a field and $p_1(x), \ldots, p_r(x)$ distinct, monic, irreducible polynomials in $F[x]$. Let $f(x) = p_1(x)^{n_1} + \cdots + p_r(x)^{n_r}$ where each $n_i$ is a positive integer.

		\begin{enumerate}[(a)]
			\item Determine the number of ideals in $F[x]/\cyc{f(x)}$.
			\item Determine the number of prime ideals in $F[x]/\cyc{f(x)}$.
		\end{enumerate}
	\end{prob}

	\begin{prob}{F13}{F13.LA1}
		Let $T: \R^3 \to \R^3$ be a lienar transformation. Prove the following:
		\begin{enumerate}[(a)]
			\item $T$ has a one-dimensional invariant subspace.
			\item $T$ has a two-dimensional invariant subspace.
		\end{enumerate}

		\emph{ Do not use the existence of canonical forms for matrices.}
	\end{prob}

	\begin{prob}{F13}{F13.LA2}
		Let $V$ be a finite dimensional vector space of $\Q$ and let $M$ be an automorphism of $V$ such that $M$ fixes no non-zero vector in $V$. Suppose that $M^p$ is the identyt map on $V$ with $p$ a prime. Show that the dimension of $V$ is divisible by $p-1$. \emph{ You may assume that the polynomial $x^{p-1} + \cdots + x + 1$ is irreducible  over $\Q$.}
	\end{prob}

	\begin{prob}{F13}{F13.LA3}
		Let $\R^3$ have  the usual inner product and suppose that $(a,b,c) \in \R^3$ is of unit length. Let $W$ be the plane given by $ax + by + cz = 0$. Let $\ell$ be the line through the origin in the direction $(a,b,c)$.
		\begin{enumerate}[(a)]
			\item Define the standard matrix representing the orthogonal projection of $\R^3$ onto $\ell$.
			\item Define the standard matrix representing the orthogonal projection of $\R^3$ onto $W$.
		\end{enumerate}
	\end{prob}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	

	\begin{prob}{F19}{}
	Let $p$ be a prime number and suppose that $1 \leq n <p^2$ is an integer. Show that every Sylow $p$-subgroup of the symmetric group $S_n$ is abelian.
	\end{prob}
	

	\begin{prob}{F20}{F20.G3}
	Let $G$ be a finite group, $x \in G$, and $H$ a subgroup of $G$.
	\begin{enumerate}[(a)]
	\item Prove that the number of conjugates of $x$ in $G$ divides $\rm{order}(G)/\rm{order}(x)$.
	\item Prove that the number of conjugates of $H$ in $G$ divides the index of $H$ in $G$.
	\end{enumerate}
	\end{prob}

	\begin{prob}{S20}{}
	Let $K$ be a field. Show that every finite subgroup of $K^\times$ is cyclic.
	\end{prob}

	\begin{prob}{S19}{}
	State and prove Lagrange's Theorem. Prove that a subgroup of a cyclic group is cyclic.
	\end{prob}

	\begin{prob}{S19}{}
	Prove the following:
	\begin{enumerate}[(a)]
	\item If $R$ is a commutative ring with no nilpotent elements, then $R[x]$ has no nilpotent elements.
	\item If $r$ is a nilpotent element of a ring with unity, then $1 - r$ is a unit.
	\end{enumerate}
	\end{prob}
	

	
	\begin{prob}{S19}{}
	Prove that $\sigma,\tau \in S_n$ are conjugate if and only if for each $m \geq 2$ the number of $m$-cycles in a cycle decomposition of $\sigma$ equals the number of $m$-cycles in a cycle decomposition of $\tau$.
	\end{prob}
	
	\begin{prob}{S20}{}
	\begin{enumerate}[(a)]
	\item Prove that the centralizer of an element is a subgroup.
	\item If $G$ is a finite group, prove that the number of elements in the conjugacy class divides the order of $G$.
	\end{enumerate}
	\end{prob}
	
	\begin{prob}{F20}{F20.G1}
	Show that any finitely generated group of $(\Q,+)$ is cyclic. Use this to prove that the direct product $(\Q,+) \times (\Q,+)$ and $(\Q,+)$ are not isomorphic.
	\end{prob}

	\begin{prob}{S20}{}
	Let $R$ be a commutative ring with identity. Show that if $p$ is a prime ideal in $R$ then
		\[p(x) = \left\{\textrm{all polynomials } \sum_i a_ix^i \textrm{ with each } a_i \in p \right\} \]
	is a prime ideal in $R[x]$.
	\end{prob}
	
	\begin{prob}{S20}{}
	Let $\F_3$ be the field with 3 elements.
	\begin{enumerate}[(a)]
	\item Prove that $K = \F_3[x]/(x^2 + 1)$ is a field.
	\item How many elements does $K$ have?
	\item Prove that $x+1$ generates the multiplicative group of non-zero elements in $K$.
	\end{enumerate}
	\end{prob}

	
	\begin{prob}{F20}{F20.RF1}
	The polynomial $x^3 - x$ has six roots in the ring $\Z/6\Z$. Find a sufficient condition on a commutative ring $R$ which ensures that the number of roots of a polynomial with coefficients in $R$ cannot exceed its degree and justify your assertion.
	\end{prob}
	
	\begin{prob}{F20}{F20.RF2}
	Prove or provide a counter example: Suppose that $K$ is a finite extension of $F$. $F \sq L \sq K$, $F \sq M \sq KL$, $LM = K$ and $L \cap M = F$. Then $[L:F][M:F] = [K:F]$. Here, $LM$ denotes the composition of the fields $L$ and $M$.
	\end{prob}
	
	\begin{prob}{F20}{F20.RF3}
		\begin{enumerate}[(i)]
		\item Let $R$ be a UFD and $d$ a nonzero element in $R$. Prove that there are only finitely many principal ideals  in $R$ that contain the ideal $(d)$.
		\item Give an example of a UFD $R$ and a nonzero element $d \in R$ such that there are infinitely many ideals in $R$ that contain $(d)$.
		\end{enumerate}
	\end{prob}
	
	\begin{prob}{F20}{F20.LA1}
	It is known that real symmetric matrices are always diagonalizable. You may assume this fact.
	\begin{enumerate}[(a)]
	\item What special property do the eigenspaces of a real symmetric matrix have?
	\item Prove that any real symmetric matrix $S$ can be diagonalized by an orthonormal matrix $U$.
	\end{enumerate}
	\end{prob}
	
	\begin{prob}{F20}{F20.LA2}
	Prove or give a counter example.
	\begin{enumerate}[(a)]
	\item If a $4 \times 4$ real matrix has characteristic polynomial $x^4 -1$ then its minimal polynomial cannot be $x^2 - 1$.
	\item Every $n \times n$ real matrix is similar over the reals to an upper triangular matrix.
	\end{enumerate}
	\end{prob}
	
	\begin{prob}{F20}{F20.LA3}
	Let $V$ be a finite dimensional complex vector space and $T: V \to V$ a linear transformation.
	\begin{enumerate}[(a)]
	\item Show that $V$ has a "flag" of subspaces $V_0 = 0 \sq V_1 \sq \cdots \sq V_n = V$ such that $\dim(V_i) = i$ and $T(V_i) \sq V_i$ for each $i$.
	\item Show that there is a basis for $V$ such that the matrix of $T$ with respect to this basis is upper triangular.
	\end{enumerate}
	\end{prob}
	

	
	\begin{prob}{F21}{F21.G3}
	Let $p$ and $q$ be distinct odd primes. Use the Sylow Theorems to show that every group of order $p^2q^2$ is not simple.
	\end{prob}	
	
	\begin{proof}
	Suppose that $|G| = p^2q^2$ and without loss of generality, assume that $p < q$. Note that each Sylow $p$-subgroup is of order $p^2$ and each Sylow $q$-subgroup is of order $q^2$. Furthermore, the number of Sylow-$p$ subgroups is $n_p = 1, q, q^2$ and the number of Sylow-$q$ subgroups is $n_q = 1, p^2$. If $n_q = 1$, then the Sylow-$q$ subgroup is normal in $G$ and thus $G$ is not simple. Suppose instead that $n_q = p^2$.\\
	
	If $Q_i, Q_j$ are any two distinct Sylow-$q$ subgroups, the intersection $Q_i \cap Q_j$ is either of size $1$ or $q$. If every pairwise intersection between the Sylow-$q$ subgroups is trivial, then the $p^2$ Sylow-$q$ subgroups account for $p^2q^2 - (p^2 -1)$ elements implying that $n_p = 1$. That is, the Sylow-$p$ subgroup is normal and therefore $G$ is not simple. Suppose that some Sylow-$q$ subgroups$Q_1, Q_2$ have intersection of size $q$. Let $N = Q_1 \cap Q_2$ and let $M$ be the subgroup of $G$ generated by $Q_1$ and $Q_2$. Then, $|M| > |Q_1| = q^2$ and $|M|$ divides $|G| = p^2q^2$. Therefore $|M| = pq^2$ or $|M| = p^2 q^2$. 
	
Note first that any group of prime squared order is abelian. Therefore both $Q_1$ and $Q_2$ are abelian and so $N = Q_1 \cap Q_2$ is normal in $M$. If $|M| = p^2q^2$ then $M=G$ and thus $N$ is a normal subgroup of $G$ and is nontrivial.\\

Now suppose that $|M| = pq^2$. 
	\end{proof}
	
	\begin{prob}{S19}{}
	\begin{enumerate}[(a)]
	\item Suppose that $G$ is a group and $G/Z(G)$ is cyclic. Prove  that $G$ is abelian.
	\item Let $p$ be a prime number and $G$ a non-cyclic finite $p$-group. Prove that $G$ contains a normal subgroup $N$ such that $G/N \cong C \oplus C$ where $C$ is a cyclic group of order $p$.
	\end{enumerate}
	\end{prob}
	
	\begin{proof}
	See \ref{prob:GZG}.
	\end{proof}
	
	\begin{proof}
	Suppose that $|G| = p^k$ with $k \geq 2$. Assume that $G$ is not cyclic.
	\end{proof}
	
\end{document}