\documentclass[../../AlgebraQualSolutions.tex]{subfiles}

\begin{document}

\subsection{Nilpotent}

\begin{prob}{S19}{}
	Prove the following:
	\begin{enumerate}[(a)]
	\item If $R$ is a commutative ring with no nilpotent elements, then $R[x]$ has no nilpotent elements.
	\item If $r$ is a nilpotent element of a ring with unity, then $1 - r$ is a unit.
	\end{enumerate}
	\end{prob}
	
	\begin{proof}
		Suppose that $a_0 + a_1 x + \cdots + a_nx^n = p(x) \in R[x]$ is nilpotent. Then there exists $k \in \N$ such that $p(x)^k = (a_0 + a_1 x + \cdots + a_nx^n)^k = 0$. In particular, the leading coefficient of $p(x)^k$ is $a_n^k$ and since $p(x)^k = 0$, $a_n^k = 0$. Since $a_n$ is the leading term of $p(x)$, it is nonzero and therefore a nilpotent element of $R$.
	\end{proof}
	
	\begin{proof}
		Suppose that $r$ is nilpotent. Choose a minimal $k \in \N$ such that $r^k = 0$. Then,
			\[1 = 1 - r^k = (1-r)(1 + r + r^2 + \cdots + r^{k-1})\]
		meaning that $1-r$ has multiplicative inverse given by $(1 + r + r^2 + \cdots + r^{k-1})$. That is, $1-r$ is a unit.
	\end{proof}

\begin{prob}{F14.RF3}{F14.RF3}
	Prove that the set $N$ of nilpotent elements of a commutative ring $R$ is an ideal of $R$ and that $R/N$ has no nonzero nilpotent elements.
\end{prob}

\begin{proof}
	Suppose that $x,y \in N$. Let $m,n \in \N$ be such that $x^m = 0$ and $y^n = 0$. Then,
		\[(xy)^m = x^my^m = 0 \]
	meaning that $xy \in N$. Since $R$ is commutative, the Binomial Expansion Theorem holds and so
		\[(x + y)^{m+n} = \sum_{k=0}^{m+n} {{m+n}\choose{k}}x^ky^{m+n - k}.\]
	Observe that whenever $k \geq m$, $x^k = 0$ and whenever $k \geq n$, $y^k = 0$. When $0 \leq k \leq m$, $m + n - k \geq n$ and therefore, ${{m+n}\choose{k}}x^ky^{m+n - k} = 0$. When $m \leq k \leq m+n$, ${{m+n}\choose{k}}x^ky^{m+n - k} = 0$. Therefore, $(x+y)^{m+n} = 0$ meaning that $x + y \in N$.\\

	Assume now that $r \in R$ is arbitrary. Then,
		\[(rx)^m = r^mx^m = 0\]
	and so $rx \in N$. As $R$ is a commutative ring, this proves that $N$ is an ideal of $R$.
\end{proof}

\begin{proof}
	Suppose that $r + N$ is nilpotent in $R/N$. Choose $m \in \N$ such that $(r + N)^m = N$. For this $m$, it follows that $r^m + N = N$ or equivalently, $r^m \in N$. Since $r^m \in N$, there exists $n \in \N$ where $(r^m)^n = 0$. But this means that $r^{mn} = 0$ and so $r \in N$. Thus, any nilpotent element of $R/N$ is zero.
\end{proof}

\begin{prob}{(F12.RF3)}{F12.RF3}
\begin{enumerate}[(a)]
\item Prove that $\Z/m\Z$ has no non-zero nilpotent elements if and only if $m$ has no multiple prime factor.
\item Prove that every element of $\Z/m\Z$ is either nilpotent or a unit whenever $m$ is a prime power.
\item Prove that if $r$ is a nilpotent element of a ring with unity then $1-r$ is a unit.
\end{enumerate}
\end{prob}

\begin{proof}
    Suppose first that $m$ has some multiple prime factor. If $p_1,\ldots, p_n$ are the distinct prime factors of $m$, this implies that there exists $k \geq 2$ for which some $p_j^k | m$. In particular, this means that the product $p_1 \cdots p_n$ is not congruent to zero modulo $m$. Write out the prime factorization for $m$ as 
		\[m = p_1^{\ell_1} \cdots p_n^{\ell_n}\]
	and define $d = \gcd{\ell_1,\ldots, \ell_n}$. Then, $(p_1\cdots p_n)^d = 0$ in $\Z/m\Z$ and thus is a nonzero nilpotent element.\\

	Now suppose that $m = p_1 \cdots p_n$ with $p_1,\ldots, p_n$ the distinct prime factors of $m$. That is, $m$ has no multiple prime factors. Let $x \in \Z/m\Z$ be such that $x^k = 0$ for some $k \in \N$. Write $x = q_1^{\ell_1} \cdots q_r^{\ell_r}$ where each $q_j$ is prime, the $q_j$ are distinct, and $0 \leq x < m$. Then, $x^k = q_1^{k\ell_1} \cdots q_r^{k\ell_r}$. Since $m$ divides $x^k$, each $p_j$ must divide $x^k$. Since the $q_i$ and $p_j$ are all prime, this means that $p_j = q_i$ for some $i$. But, $x$ and $x^k$ share the same prime factors and so each $p_j$ must divide $x$. Therefore, $m$ divides $x$ and so $x = 0$ in $\Z/m\Z$. It follows that any nilpotent element in such a $\Z/m\Z$ must be zero.
\end{proof}

\begin{proof}
    Let $p$ be prime and suppose that $m = p^N$ for some $N \in \N$. Let $x \in \Z/m\Z$ and suppose first that $p$ divides $x$. Then $x = pq$ for some $q$. Observe that
        \[x^N = (pq)^N = p^Nq^N = 0\]
    since $p^N = m \equiv 0$ in $\Z/m\Z$. That is, whenever $p$ divides $x$, it follows that $x$ is nilpotent. Suppose now that $p$ does not divide $x$. Because the only divisors of $m$ are powers of $p$, the greatest common divisor of $m$ and $x$ is 1. Therefore there exist integers $s,t$ where $1 = xs + mt$. As $mt = 0$ in $\Z/m\Z$, $xs = 1$ implying that $x$ is a unit.
\end{proof}

\begin{proof}
    Assume that $r \in R$ is a nilpotent element of a ring with unity. Then there exists $n \in \N$ such that $r^n = 0$. Then,
        \[1 = 1 - r^n = (1 + r + r^2 + \cdots + r^{n-1})(1-r)\]
    meaning that $\sum_{i=0}^{n-1}r^i$ is the multiplicative inverse of $(1-r)$. That is, $1-r$ is a unit.
\end{proof}

\end{document}