\documentclass[../AlgebraQualSolutions.tex]{subfiles}

\begin{document}

\subsection{Characteristic and Minimal Polynomials}
 

\begin{prob}{(F04.L1)}{F04.L1}
	Let $V$ be a four-dimensional vector space over $\R$ with basis $\{v_1,v_2,v_3,v_4\}$ and let $T: V \to V$ be a linear transformation such that
		\[T(v_1) = v_2 \quad T(v_2) = v_3 \quad T(v_3) = v_4 \quad T(v_4) = -v_1 -4v_2 -6v_3 - 4fv_4.\]
	\begin{enumerate}[(a)]
		\item What is the characteristic polynomial of $T$?
		\item Is $T$ diagonalizable over $\C$?
	\end{enumerate}
\end{prob}

\begin{solution}
	Consider the matrix representation of $T$ with respect to the given basis:
		\[\begin{pmatrix}
			0 & 0 & 0 & -1\\
			1 & 0 & 0 & -4\\
			0 & 1 & 0 & -6\\
			0 & 0 & 1 & -4 \end{pmatrix}.\]
		Observe that with respect to this basis, the matrix of $T$ is in rational canonical form with one $4 \times 4$ block. From the last column of this matrix, it follows that the minimal polynomial of $T$ is 
			\[m(x) = x^4 + 4x^3 + 6x^2 + 4x + 1.\]
		The characteristic polynomial $c(x)$ of $T$ is a monic, degree-4 polynomial. Furthermore, $m$ must divide $c$ and $m$ and $c$ have the same roots (disregarding multiplicity). This implies that $c(x) = m(x) = (x + 1)^4$
\end{solution}

\begin{claim}
	$T$ is not diagonalizable over $\C$.

	\begin{proof}
		The minimal and characteristic polynomials of $T$ are both $(x + 1)^4$. Since the minimal polynomial has repeated roots, $T$ is not diagonalizable.
	\end{proof}
\end{claim}

\begin{prob}{(F20.LA2)}{F20.LA2}
	Prove or give a counter example.
	\begin{enumerate}[(a)]
	\item If a $4 \times 4$ real matrix has characteristic polynomial $x^4 -1$ then its minimal polynomial cannot be $x^2 - 1$.
	\item Every $n \times n$ real matrix is similar over the reals to an upper triangular matrix.
	\end{enumerate}
	\end{prob}
	
	\begin{proof}
		Notice that $x^4 - 1$ has four distinct roots, namely $\pm 1, \pm i$. Since $x^2 - 1$ has roots $\pm 1$ and the characteristic polynomial divides some power of the minimal polynomial, it is impossible for $x^2 -1$ to be the minimal polynomial.
	\end{proof}
	
	\begin{solution}
		This is false: consider the matrix $A = \begin{pmatrix} 0 & -1\\ 1 & 0\end{pmatrix}$. This matrix has no eigenvalues over the reals. If $A$ were to be similar to an upper triangular matrix, the (1,1) entry would need to correspond to an eigenvalue which is impossible.
	\end{solution}

\begin{prob}{(F14.LA2)}{F14.LA2}
	Prove or provide a counterexample: the characteristic polynomial of a matrix with entries in a field $K$ must be irreducible in the ring $K[x]$.
\end{prob}

\begin{solution}
	This is false. Let $K = \R$ and consider the matrix
		\[A = \begin{pmatrix} 1 & 0 \\ 0 & 1 \end{pmatrix}. \]
	The characteristic polynomial of $A$ is $(x-1)^2$ which is clearly reducible over $\R[x]$.
\end{solution}

\end{document}